---
title: "Exam_2"
format: html
---

## 2 Background


```{r}
library(tidytuesdayR)
library(tidyverse)
library(rnaturalearth)
library(sf)
```

```{r}
tuesdata <- tt_load('2020-02-18')
fc <- tuesdata$food_consumption
```
```{r}
str(fc)
head(fc, 22)
```
```{r}
fc %>%
  count(food_category)
```
What is the problem with these values?
The length of the names

List the necessary wrangling function to shortne the food category names shown in the table above. The resulting dataframe should be stored to a variable called fcc
1. fct_recode()
2. summarize()
3. str_replace_all()
4. str_replace()
```{r}
fc_cleaned <- fc %>%
  mutate(food_category = case_when(
    food_category == "Lamb & Goat" ~ "Lamb",
    food_category == "Milk - inc. cheese" ~ "Dairy",
    food_category == "Wheat and Wheat Products" ~ "Wheat",
    food_category == "Nuts inc. Peanut Butter" ~ "Nuts",
    TRUE ~ food_category   # keep everything else the same
  ))
```

```{r}
fc_cleaned %>%
  count(food_category)

```

## 3
Research question 1:Which 5 countries consume the most food are?

List the neccessary steo to produce the needed dataframe:
1. select(country,consumption)
2. group_by(country)
3. summarize(consumption=sum(consumption))
4. arrange
5.head

which graph types can be used to visualize the produced dataframe?
1. boxplots
2. bar graphs
3 bar chart
4. geom column

```{r}
library(dplyr)

top5_countries <- fc %>%
  group_by(country) %>%
  summarise(total_consumption = sum(consumption, na.rm = TRUE)) %>%
  arrange(desc(total_consumption)) %>%
  slice_head(n = 5)
print(top5_countries)


```

```{r}
library(ggplot2)

ggplot(top5_countries, aes(x = reorder(country, total_consumption), y = total_consumption)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(title = "Top 5 Countries by Total Food Consumption",
       x = "Country",
       y = "Total Consumption (kg per person per year)") +
  theme_minimal()

```

## 4
Research Question 2: Which top 5 countries consume each food are?

List the neccessary step to compute the required dataframe
1. select
2. group_by
3. slice_max

which graphs can be used to visualize this data frame?
1. facet
2. columns

```{r}
# Step 1: Get top 5 countries by food category
top5_by_food <- fc_cleaned %>%
  group_by(food_category) %>%
  arrange(desc(consumption)) %>%    # sort within each food
  slice_head(n = 5) %>%              # take top 5 per food
  ungroup()

# Step 2: View the 3-column table
top5_by_food %>%
  select(food_category, country, consumption)
```
```{r}
# Plot: top 5 countries for each food
ggplot(top5_by_food, aes(x = reorder(country, consumption), y = consumption, fill = food_category)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  facet_wrap(~ food_category, scales = "free_y") +  # one small plot per food
  labs(
    title = "Top 5 Countries by Food Consumption for Each Food Category",
    x = "Country",
    y = "Consumption (kg per person per year)"
  ) +
  theme_minimal()

```
## 5
Research question 3: What does the consumption of each food look like?



```{r}
ne_countries(returnclass = "sf") |>
  select(name, geometry) |>
  left_join(fc_cleaned |> select(-co2_emmission), join_by(name == country)) |>
  ggplot() + geom_sf(aes(fill = consumption)) + facet_wrap(~food_category) + theme(legend.position = "bottom")

```
The plot above sufferes from a bad color scheme, no units, no boundaries, no title

```{r}
library(ggplot2)
library(dplyr)
library(sf)

ne_countries(returnclass = "sf") |>
  select(name, geometry) |>
  left_join(fc_cleaned |> select(-co2_emmission), by = c("name" = "country")) |>
  ggplot() + 
  geom_sf(aes(fill = consumption), color = "white") +  # Color borders for countries
  facet_wrap(~food_category) + 
  scale_fill_viridis_c(option = "plasma", name = "Consumption (kg per person per year)") + # Better color scheme
  labs(
    title = "Global Food Consumption by Category",
    subtitle = "Top 5 Countries by Food Consumption for Each Category",
    caption = "by PJ Pozniak"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    strip.text = element_text(size = 12, face = "bold"),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 12, hjust = 0.5),
    plot.caption = element_text(size = 10, hjust = 1)
  )

```
This one has a better color scheme, a title, and better labels

```{r}
library(dplyr)

# Calculate the min, max, and range for each food category
food_summary <- fc_cleaned |>
  group_by(food_category) |>
  summarise(
    min_consumption = min(consumption, na.rm = TRUE),
    max_consumption = max(consumption, na.rm = TRUE),
    range_consumption = max(consumption, na.rm = TRUE) - min(consumption, na.rm = TRUE)
  )

# Print the summary
print(food_summary)

```

