---
title: "Spatial Viz"
---

Use this file for practice with the **spatial viz** in-class activity.  Refer to the class website for details.

```{r}
fave_places <- read.csv("https://hash-mac.github.io/stat112site-s25/data/our_fave_places.csv")

# Check it out
head(fave_places)
```
## Exercise 1

```{r}
# Load the leaflet package
install.packages("leaflet")
library(leaflet)

# Just a plotting frame
leaflet(data = fave_places)
```

```{r}
# Now what do we have?
leaflet(data = fave_places) |> 
  addTiles()
```

```{r}
# Now what do we have?
# longitude and latitude refer to the variables in our data
leaflet(data = fave_places) |> 
  addTiles() |> 
  addMarkers(lng = ~longitude, lat = ~latitude)
```

```{r}
# Since we named them "longitude" and "latitude", the function
# automatically recognizes these variables. No need to write them!
leaflet(data = fave_places) |> 
  addTiles() |> 
  addMarkers()
```

## Exercise 2
```{r}
# Load package needed to change color
library(ggplot2)

# We can add colored circles instead of markers at each location
leaflet(data = fave_places) |> 
  addTiles() |> 
  addCircles(color =("red"))
```

```{r}
# We can change the background
# Mark locations with yellow dots
# And connect the dots, in their order in the dataset, with green lines
# (These green lines don't mean anything here, but would if this were somebody's travel path!)
leaflet(data = fave_places) |>
  addProviderTiles("USGS") |>
  addCircles(weight = 10, opacity = 1, color = ("yellow")) |>
  addPolylines(
    lng = ~longitude,
    lat = ~latitude,
    color = ("green")
  )
```

## Exercise 3

```{r}
# Import starbucks location data
starbucks <- read.csv("https://mac-stat.github.io/data/starbucks.csv")
```

```{r}
# Don't worry about the syntax
starbucks_mn <- starbucks |>   
  filter(Country == "US", State.Province == "MN")
```

```{r}
leaflet(data = starbucks_mn) |> 
  addTiles() |> 
  addMarkers()
```

```{r}
ggplot(starbucks, aes(y = Latitude, x = Longitude)) + 
  geom_point(size = 0.5)
```

## Exercise 4

```{r}
# Load the package
install.packages("rnaturalearth")
library(rnaturalearth)

# Get info about country boundaries across the world
# in a "sf" or simple feature format
world_boundaries <- ne_countries(returnclass = "sf")
```

```{r}
# This code creates a world map of all the country boundaries
# We are using geom_sf
ggplot(world_boundaries) + 
  geom_sf()
```

```{r}
# Load package needed to change map theme
library(mosaic)

# Add a point for each Starbucks
# NOTE: The Starbucks info is in our starbucks data, not world_boundaries
# How does this change how we use geom_point?!
ggplot(world_boundaries) + 
  geom_sf() + 
  geom_point(
    data = starbucks,
    aes(x = Longitude, y = Latitude),
    alpha = 0.3, size = 0.2, color = "darkgreen"
  ) +
  theme_map()
```
I learned that the most amount of starbucks are in North America With other clumps in eastern europe and western Asia. There are very little starbucks in south america, eastern Asia, africa, and australia

## Exercise 5
```{r}
# We'll learn this syntax soon! Don't worry about it now.
starbucks_cma <- starbucks |> 
  filter(Country %in% c('CA', 'MX', 'US'))
```



```{r}
cma_boundaries <- ne_states(
  country = c("canada", "mexico", "united states of america"),
  returnclass = "sf")
```



```{r}
# Just the boundaries
library(tidyverse)
ggplot(cma_boundaries) + 
  geom_sf()
```

## Exercise 6
```{r}
starbucks_midwest <- starbucks |> 
  filter(State.Province %in% c("MN", "ND", "SD", "WI"))
```

```{r}
# Load packages
library(sf)
library(maps)

# Get the boundaries
midwest_boundaries <- st_as_sf(
  maps::map("county",
            region = c("minnesota", "wisconsin", "north dakota", "south dakota"), 
            fill = TRUE, plot = FALSE))

# Check it out
head(midwest_boundaries)
```

```{r}
ggplot(midwest_boundaries) +
  geom_sf() +
  geom_point(
    data = starbucks_midwest,
    aes(x = Longitude, y = Latitude),
    alpha = 0.7,
    size = 0.2,
    color = 'darkgreen'
  ) +
  theme_map()
```
## Exercise 7
```{r}
# Point map (we made this earlier)
ggplot(cma_boundaries) + 
  geom_sf() + 
  geom_point(
    data = starbucks_cma,
    aes(x = Longitude, y = Latitude),
    alpha = 0.3,
    size = 0.2,
    color = "darkgreen"
  ) +
  coord_sf(xlim = c(-179.14, -50), ylim = c(14.54, 83.11)) +
  theme_map()
```

```{r}
# What changed in the plot?
# What changed in our code?!
ggplot(cma_boundaries) + 
  geom_sf() + 
  geom_density_2d(
    data = starbucks_cma,
    aes(x = Longitude, y = Latitude),
    size = 0.2,
    color = "darkgreen"
  ) +
  coord_sf(xlim = c(-179.14, -50), ylim = c(14.54, 83.11)) +
  theme_map()
```

```{r}
elections_by_state <-  read.csv("https://mac-stat.github.io/data/election_2020_by_state.csv")
elections_by_counties <- read.csv("https://mac-stat.github.io/data/election_2020_county.csv")
```

```{r}
# Don't worry about the code!

elections_by_state <- elections_by_state |> 
  filter(state_abbr != "DC") |> 
  select(state_name, state_abbr, repub_pct_20) |> 
  mutate(repub_20_categories = 
           cut(repub_pct_20, 
               breaks = seq(30, 70, by = 5), 
               labels = c("30-34", "35-39", "40-44", "45-49",
                          "50-54", "55-59", "60-64", "65-70"), 
               include.lowest = TRUE))

elections_by_counties <- elections_by_counties |> 
  select(state_name, state_abbr, county_name, county_fips,
          repub_pct_20, median_age, median_rent) |> 
  mutate(repub_20_categories = 
           cut(repub_pct_20, 
               breaks = seq(0, 100, by = 10),
               labels = c("0-9", "10-19", "20-29", "30-39", "40-49",
                          "50-59", "60-69", "70-79", "80-89", "90-100"),
               include.lowest = TRUE))
```

## Exercise 8
```{r}
# Get the latitude and longitude coordinates of state boundaries
states_map <- map_data("state")

# Check it out
head(states_map)
```

```{r}
# Note where the dataset, elections_by_state, is used
# Note where the background map, states_map, is used
ggplot(elections_by_state, aes(map_id = state_name, fill = repub_pct_20)) +
  geom_map(map = states_map) +
  expand_limits(x = states_map$long, y = states_map$lat) +
  theme_map() 
```
```{r}
# Make it nicer!
ggplot(elections_by_state, aes(map_id = state_name, fill = repub_pct_20)) +
  geom_map(map = states_map) +
  expand_limits(x = states_map$long, y = states_map$lat) +
  theme_map() + 
  scale_fill_gradientn(name = "% Republican", colors = c("blue", "purple", "red"), values = scales::rescale(seq(0, 100, by = 5)))
```

```{r}
ggplot(elections_by_state, aes(map_id = state_name, fill = repub_20_categories)) +
  geom_map(map = states_map) +
  expand_limits(x = states_map$long, y = states_map$lat) +
  theme_map()
```

```{r}
# Load package needed for refining color palette
library(RColorBrewer)

# Now fix the colors
ggplot(elections_by_state, aes(map_id = state_name, fill = repub_20_categories)) +
  geom_map(map = states_map) +
  expand_limits(x = states_map$long, y = states_map$lat) +
  theme_map() + 
  scale_fill_manual(values = rev(brewer.pal(8, "RdBu")), name = "% Republican")
```

```{r}
# Get only the starbucks data from the US
starbucks_us <- starbucks |> 
  filter(Country == "US")

# Map it
ggplot(elections_by_state, aes(map_id = state_name, fill = repub_20_categories)) +
  geom_map(map = states_map) +
  geom_point(
    data = starbucks_us,
    aes(x = Longitude, y = Latitude),
    size = 0.05,
    alpha = 0.2,
    inherit.aes = FALSE
  ) +
  expand_limits(x = states_map$long, y = states_map$lat) +
  theme_map() + 
  scale_fill_manual(values = rev(brewer.pal(8, "RdBu")), name = "% Republican")
```

